---
title: "Coefficient Of Extremal Asymmetry"
author: "Mongi Nouira"
output: github_document
---

We introduce a coefficient of extremal asymmetry which quantifies the level of asymmetry in the structure of dependence between two random variables at extreme levels.

```{r, include=FALSE}
library(htmltools)

knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})

rmd_filename <- stringr::str_remove(knitr::current_input(), "\\.Rmd")
knitr::opts_chunk$set(
  fig.path = stringr::str_c(file.path("cache", rmd_filename), "/")
)

knitr::opts_chunk$set(
  echo=FALSE, cache=FALSE, imgcenter=TRUE, 
  fig.align="center", dpi=150, fig.width=7
)

library("evd")

# Color Map
c1 = "#636EFA"
c2 = "#EF553B"
c3 = "#00CC96"
```

```{r}
df = read.csv("~/Documents/GitHub/PDM_2022/data/data1.csv")

ll = -100*sapply(df[seq(2,ncol(df))], function(x) log(x[-1])-log(x[-length(df$Date)]))
ll = data.frame(ll)

# Sectors
sectors = sapply(colnames(ll)[1:12], function(x) substring(x,5))
sectors = as.vector(sectors)
```

## Extremal Symmetry - Examples

The number $\rho$ reflects the tendency of one variable to be more extreme than the other, given that both variables are extreme. We are interested in the limit.

```{r}
rho = function(c,q){
  n = nrow(c)
  sc = apply(c,2,rank)/(n+1)
  u = sc[,1]
  v = sc[,2]
  r = (sum(v>u & u>q) - sum(v<u & v>q)) / (sum(v>u & u>q) + sum(v<u & v>q))
  r
}
```

```{r}
s1 = 1
s2 = 4

q = seq(0.01,0.99,0.01)
rho_q = sapply(q, function(t) rho(ll[,c(s1,s2)],t))

plot(
  q, rho_q, type="l", xlab="q", ylab="rho",
  xlim=c(0,1), ylim=c(-0.5,0.5),
  main=paste(sectors[s1],sectors[s2],sep="/")
)
abline(h=0, col=c2, lty=3)
```
```{r}
s1 = 11
s2 = 12

q = seq(0.01,0.99,0.01)
rho_q = sapply(q, function(t) rho(ll[,c(s1,s2)],t))

plot(
  q, rho_q, type="l", xlab="q", ylab="rho",
  xlim=c(0,1), ylim=c(-0.5,0.5),
  main=paste(sectors[s1],sectors[s2],sep="/")
)
abline(h=0, col=c2, lty=3)
```
## Extremal Symmetry - All Sector Pairs Matrix

We look at the value for the 95% quantile which simply means using the threshold 0.95 to compute the coefficient.

```{r}
q = 0.95

phi_mx = matrix(0, nrow=12, ncol=12, dimnames=list(sectors,sectors))

for (i in seq(1,12)){
  for (j in seq(1,12)){
    
    if (i!=j){
      phi = rho(ll[,c(i,j)],q)
    } else{
      phi = NA
    }
    
    phi_mx[i,j] = phi
    
  }
}
```

```{r message=FALSE, warning=FALSE}
library("gplots")
library("RColorBrewer")
pal = brewer.pal(10,"RdYlGn")

heatmap.2(
  phi_mx, Rowv=NA, Colv=NA, dendrogram="none",
  col=pal, density.info="none", trace="none",
  cexCol=1, cexRow=1, srtCol=45, offsetCol=0,
  margins=c(8,15), keysize=1.5,
  labCol=sectors, labRow=sectors
)
```

