---
title: "Multivariate Tail Dependence Decomposition"
author: "Mongi Nouira"
output: github_document
---

We propose a decomposition which help to summarize and describe multivariate tail dependence. It can be seen as an adaptation of PCA to the extremes.

```{r, include=FALSE}
library(htmltools)

knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})

rmd_filename = stringr::str_remove(knitr::current_input(), "\\.Rmd")
knitr::opts_chunk$set(
  fig.path = stringr::str_c(file.path("cache", rmd_filename), "/")
)

knitr::opts_chunk$set(
  echo=FALSE, cache=FALSE, imgcenter=TRUE,
  fig.align="center", dpi=150, fig.width=8, fig.height=6
)

# Color Map
c1 = "#636EFA"
c2 = "#EF553B"
c3 = "#00CC96"

library("ismev")
library("evir")
library("gplots")
library("lattice")
library("RColorBrewer")
```

```{r}
df = read.csv("~/Documents/GitHub/PDM_2022/data/data1.csv")

# Data
ll = -100*sapply(df[seq(2,ncol(df))], function(x) log(x[-1])-log(x[-length(df$Date)]))
ll = data.frame(ll)

# Sectors
sectors = sapply(colnames(ll)[1:12], function(x) substring(x,5))
sectors = as.vector(sectors)
colnames(ll) = sectors
```

## Data Transformation

1. Keep only positive losses.
2. Apply bijection x->log(1+exp(x)).
3. Scale variables using empirical/gpd mixture to Frechet scale transformation.

The last two transformations are required by the framework. In particular, the tail index needs to be 2.

```{r}
# Plot 1
par(pty="s")
pairs(ll[,seq(1,3)], upper.panel=NULL, main="Original Data")
```

```{r}
# Empirical/GPD transform
fst_correct = function(x,xe,r,sc,sh,u){
  if(x>u){-1/log(1-r*(max(1+sh*(x-u)/sc,0)^(-1/sh)))} else{xe}
}

gpd_rescale = function(x,u,emp=F,qua=F){
  
  # Scaling
  #x = x/mad(x)
  
  # Empirical transformation to Frechet scale
  xe = -1/log(rank(x)/(length(x)+1))
  
  if(emp){
    z = xe
  }else{
    if(qua){
      u = quantile(x,0.95)
    }
    # Fit gpd to exceedances and correct their transformation
    fit = gpd.fit(x, threshold=u, npy=251, show=F)
    z = mapply(
      function(a,b) fst_correct(a,b,fit$rate,fit$mle[1],fit$mle[2],u), x, xe
    )
  }
  
  # Transform unit Frechet (alpha=1) to RV alpha=2
  z = sqrt(z)
  z
}
```

```{r}
hill_rescale = function(x){
  fit = gpd.fit(x, threshold=2, npy=251, show=FALSE)
  sc = fit$mle[1]
  al = 1/fit$mle[2]
  print(paste("alpha =",round(al,2),"scale =",round(sc,2),sep=" "))
  z = sqrt((x^al)/sc)
  z
  #hill(z,end=1500)
}
```

```{r}
# Keep Positive Losses Only - Bijection
sll = apply(ll, c(1,2), function(x) log(1+exp(max(x,0))))

sll = apply(sll, 2, function(x) gpd_rescale(x,2,emp=T,qua=T))
#sll = apply(sll, 2, function(x) hill_rescale(x))
print("---")
#apply(sll, 2, function(x) hill_rescale(x))
```

```{r}
# Plot 2
par(pty="s")
pairs(sll[,seq(1,3)], log="xy", upper.panel=NULL, main="Transformed Data")
```

## Tail Pairwise Dependence Matrix

```{r}
# New Coordinates
r = apply(sll, 1, function(x) norm(as.matrix(x), type="1"))
w = sll/r

n = length(r)
r0 = quantile(r,0.95)
nr0 = sum(r>r0)

m = (r0^2)*nr0/n

W = as.matrix(w[r>r0,])

# Tail Wise Dependence Matrix
S = (m/nr0) * t(W)%*%W
```

```{r}
par(mar=c(12,4,4,4))
barplot(diag(S), las=3, ylim=c(0,1.2), main="Matrix Diagonal")
abline(h=1, lty=2, col=c2)
```

```{r}
levelplot(
  S, xlab="", ylab="", 
  col.regions=colorRampPalette(brewer.pal(8, "Blues"))(25),
  pretty=TRUE, scales=list(x=list(rot=45))
)
```

## Eigen Decomposition

```{r message=FALSE, warning=FALSE}
ED = eigen(S)

e = ED$values
u = ED$vectors

# sign change : describing losses better
u[,1] = -u[,1]
u[,2] = -u[,2]
```

### Eigen Values

```{r}
plot(e, type="l", xlab="Index", ylab="Eigen Value")
```

### Eigen Vectors

```{r}
heatmap.2(
  u[,1:4], Rowv=NA, Colv=NA, dendrogram="none",
  col=brewer.pal(10,"RdYlGn"), density.info="none", trace="none",
  cexCol=1, cexRow=1, srtCol=0,
  margins=c(5,15), keysize=2,
  labCol=c("u1","u2","u3","u4"), labRow=sectors)
```

## Principal Components Time Series

```{r}
dll = apply(ll, 1, function(x) t(u[,1:3])%*%x)
dll = t(dll)

par(mfrow=c(3,1))
plot(as.Date(df$Date[-1]), y=dll[,1], type="l", xlab="Date", ylab="v1")
plot(as.Date(df$Date[-1]), y=dll[,2], type="l", xlab="Date", ylab="v2")
plot(as.Date(df$Date[-1]), y=dll[,3], type="l", xlab="Date", ylab="v3")
```

```{r}
par(mfrow=c(3,1))
plot(as.Date(df$Date[-1]), y=dll[,1], ylim=c(min(dll),max(dll)), type="l", xlab="Date", ylab="v1")
plot(as.Date(df$Date[-1]), y=dll[,2], ylim=c(min(dll),max(dll)), type="l", xlab="Date", ylab="v2")
plot(as.Date(df$Date[-1]), y=dll[,3], ylim=c(min(dll),max(dll)), type="l", xlab="Date", ylab="v3")
```

```{r}
par(mfrow=c(1,3), pty="s")
plot(dll[,c(1,2)], xlab="v1", ylab="v2")
plot(dll[,c(1,3)], xlab="v1", ylab="v3")
plot(dll[,c(2,3)], xlab="v2", ylab="v3")
```

```{r}
par(mfrow=c(1,3), pty="s")
plot(dll[,c(1,2)], xlim=c(-5,40), ylim=c(min(dll),max(dll)), xlab="v1", ylab="v2")
plot(dll[,c(1,3)], xlim=c(-5,40), ylim=c(min(dll),max(dll)), xlab="v1", ylab="v3")
plot(dll[,c(2,3)], xlim=c(-5,40), ylim=c(min(dll),max(dll)), xlab="v2", ylab="v3")
```


