---
title: "GEV Univariate/Bivariate - EDA"
author: "Mongi Nouira"
output: github_document
---

We analyze the dependence between the technology and energy sectors.

```{r, include=FALSE}
library(htmltools)

knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})

knitr::opts_chunk$set(
  echo=FALSE, cache=FALSE, imgcenter=TRUE, 
  fig.align="center", dpi=150, fig.width=7
)

library("ismev")
library("evd")

# Color Map
c1 = "#636EFA"
c2 = "#EF553B"
c3 = "#00CC96"
```

```{r}
df = read.csv("~/Documents/GitHub/PDM_2022/data/data1.csv")

ll = -100*sapply(df[seq(3,ncol(df))], function(x) log(x[-1])-log(x[-length(df$Date)]))
#ll = apply(ll, c(1,2), function(x) max(x,0))
ll = data.frame(ll)
ll$Date = df$Date[-1]
```

## Component-Wise Block Maxima

We start by transforming the data into componentwise block maxima data by taking the maximum of groups of 25 observations. This produces 201 maxima observations.

```{r}
cml = ll[1:5025,c(2,11)]
xx = rep(1:201, each=25)
cml = cbind(tapply(cml[,1], xx, max), tapply(cml[,2], xx, max))
colnames(cml) = colnames(ll[,c(2,11)])

par(mfrow=c(1,2), pty="s")
plot(ll[,c(2,11)], main="Original And Maxima Data", xlim=c(-20,20), ylim=c(-20,20))
points(cml, col=c1)

ecml <- -1/log(apply(cml,2,rank)/(nrow(cml)+1))
plot(ecml, log="xy", main="Standard Frechet Margins")
```

## Non-Parametric

First, we use various non parametric estimates of the dependence function that also uses empirical estimation of the margins.

```{r}
#cml = ll[ll[,6]>2&ll[,7]>2,c(6,7)]

pp = "pickands"
cc = "cfg"

par(pty="s")

abvnonpar(data = cml, epmar = TRUE, method = pp, plot = TRUE, lty = 1, col=c1)
abvnonpar(data = cml, epmar = TRUE, method = pp, add = TRUE, madj = 1, lty = 2)
abvnonpar(data = cml, epmar = TRUE, method = pp, add = TRUE, madj = 2, lty = 4)
abvnonpar(data = cml, epmar = TRUE, method = cc, add = TRUE, lty = 1, col=c2)

legend(
  "bottomright", 
  legend=c("Pickands","Deheuvels","Hall-Tajvidi","Caperaa-Fougeres-Genest"), 
  col=c(c1,'black','black',c2), lty=c(1,2,4,1), cex=0.6
)

title(main="Non-Parametric Estimation Of Dependence Function")
```

```{r}
par(pty="s")
plot(ll[,c(2,11)], main="Quantile Curves", xlim=c(-20,20), ylim=c(-20,20))
points(cml, col=c1)
qcbvnonpar(c(0.95,0.99,0.995), data = cml, epmar = TRUE, mint = 30, add = TRUE, col = c2)
```

## Parametric

Second, we use maximum likelihood estimation for various parametric Bivariate Extreme Value Distributions models.

```{r}
m1 = fbvevd(cml, model = "log")
m2 = fbvevd(cml, model = "bilog")
m3 = fbvevd(cml, model = "ct")

par(pty="s")
plot(m1, which = 4, nplty = 1, col = c1, main="Parametric Estimation Of Dependence Function")
plot(m2, which = 4, nplty = 1, col = c2, add = TRUE)
plot(m3, which = 4, nplty = 1, col = c3, add = TRUE)

legend(
  "bottomright", 
  legend=c("logistic","bilogistic","Dirichlet"), 
  col=c(c1,c2,c3), lty=rep(1,3), cex=0.6
)
```

The following shows the parameter estimates and their standard errors corresponding to the log model.

```{r}
knitr::kable(round(rbind(fitted(m1), std.errors(m1)), 3))
```

Test logistic vs bilogistic model. Test of independence.

```{r}
anova(m2,m1)
evind.test(cml, method = "score")
```

