---
title: "Dependence Functions Test"
author: "Mongi Nouira"
output: github_document
---

We look at the dependence function by Pickands and other variations. We use a non-parametric estimation of dependence function.

```{r, include=FALSE}
library(htmltools)

knitr::knit_hooks$set(imgcenter = function(before, options, envir){
  if (before) {
    HTML("<p align='center'>")
  } else {
    HTML("</p>")
  }
})

rmd_filename <- stringr::str_remove(knitr::current_input(), "\\.Rmd")
knitr::opts_chunk$set(
  fig.path = stringr::str_c(file.path("cache", rmd_filename), "/")
)

knitr::opts_chunk$set(
  echo=FALSE, cache=FALSE, imgcenter=TRUE, 
  fig.align="center", dpi=150, fig.width=7
)

library("evd")
library("ismev")

# Color Map
c1 = "#636EFA"
c2 = "#EF553B"
c3 = "#00CC96"
```

```{r}
df = read.csv("~/Documents/GitHub/PDM_2022/data/data1.csv")

ll = -100*sapply(df[seq(2,ncol(df))], function(x) log(x[-1])-log(x[-length(df$Date)]))
ll = data.frame(ll)

sectors = sapply(colnames(ll)[1:ncol(ll)], function(x) substring(x,5))
sectors = as.vector(sectors)
```

## Dependence Functions Variations
We use the component wise block maxima approach.

```{r}
s1 = 1
s2 = 4

cml = ll[1:5025,c(s1,s2)]
xx = rep(1:201, each=25)
cml = cbind(tapply(cml[,1], xx, max), tapply(cml[,2], xx, max))
colnames(cml) = colnames(ll[,c(2,11)])

pp = "pickands"
cc = "cfg"

par(pty="s")

abvnonpar(data = cml, epmar = TRUE, method = pp, plot = TRUE, lty = 1, col=c1, lwd=2)
abvnonpar(data = cml, epmar = TRUE, method = pp, add = TRUE, madj = 1, lty = 2, lwd=2)
abvnonpar(data = cml, epmar = TRUE, method = pp, add = TRUE, madj = 2, lty = 4, lwd=2)
abvnonpar(data = cml, epmar = TRUE, method = cc, add = TRUE, lty = 1, col=c2, lwd=2)

legend(
  "bottomright",
  legend=c(pp,"Deheuvels","Hall-Tajvidi",cc), 
  col=c(c1,'black','black',c2), lty=c(1,2,4,1), cex=0.8
)

title(main=paste(sectors[s1],sectors[s2],sep="/"))
```

```{r}
s1 = 3
s2 = 12

cml = ll[1:5025,c(s1,s2)]
xx = rep(1:201, each=25)
cml = cbind(tapply(cml[,1], xx, max), tapply(cml[,2], xx, max))
colnames(cml) = colnames(ll[,c(2,11)])

pp = "pickands"
cc = "cfg"

par(pty="s")

abvnonpar(data = cml, epmar = TRUE, method = pp, plot = TRUE, lty = 1, col=c1, lwd=2)
abvnonpar(data = cml, epmar = TRUE, method = pp, add = TRUE, madj = 1, lty = 2, lwd=2)
abvnonpar(data = cml, epmar = TRUE, method = pp, add = TRUE, madj = 2, lty = 4, lwd=2)
abvnonpar(data = cml, epmar = TRUE, method = cc, add = TRUE, lty = 1, col=c2, lwd=2)

legend(
  "bottomright",
  legend=c(pp,"Deheuvels","Hall-Tajvidi",cc), 
  col=c(c1,'black','black',c2), lty=c(1,2,4,1), cex=0.8
)

title(main=paste(sectors[s1],sectors[s2],sep="/"))
```

## Threshold Effect
We use the threshold exceedances approach. We compute the dependence function using the observations that exceeds both thresholds.

```{r}
s1 = 1
s2 = 4

bll = ll[,c(s1,s2)]

par(pty="s")

abvnonpar(data = bll[bll[,1]>1.5 & bll[,2]>1.5,], 
          epmar=TRUE, method=cc, plot=TRUE, lty=1, col=c1, lwd=2)
abvnonpar(data = bll[bll[,1]>2 & bll[,2]>2,], 
          epmar=TRUE, method=cc, add=TRUE, lty=1, col=c2, lwd=2)
abvnonpar(data = bll[bll[,1]>2.5 & bll[,2]>2.5,], 
          epmar=TRUE, method=cc, add=TRUE, lty=1, col=c3, lwd=2)

legend(
  "bottomright",
  legend=c("1.5","2","2.5"), 
  col=c(c1,c2,c3), lty=c(1,1,1), cex=0.8
)

title(main=paste(sectors[s1],sectors[s2],sep="/"))
```

```{r}
s1 = 3
s2 = 12

bll = ll[,c(s1,s2)]

par(pty="s")

abvnonpar(data = bll[bll[,1]>1.5 & bll[,2]>1.5,], 
          epmar=TRUE, method=cc, plot=TRUE, lty=1, col=c1, lwd=2)
abvnonpar(data = bll[bll[,1]>2 & bll[,2]>2,], 
          epmar=TRUE, method=cc, add=TRUE, lty=1, col=c2, lwd=2)
abvnonpar(data = bll[bll[,1]>2.5 & bll[,2]>2.5,], 
          epmar=TRUE, method=cc, add=TRUE, lty=1, col=c3, lwd=2)

legend(
  "bottomright",
  legend=c("1.5","2","2.5"), 
  col=c(c1,c2,c3), lty=c(1,1,1), cex=0.8
)

title(main=paste(sectors[s1],sectors[s2],sep="/"))
```

## New Pickands Estimation - Test

```{r}
fst_correct = function(x,xe,r,sc,sh,u){
  if(x>u){-1/log(1-r*(max(1+sh*(x-u)/sc,0)^(-1/sh)))} else{xe}
}

gpd_rescale = function(x,u){
  # Empirical transformation to Frechet scale
  xe = -1/log(rank(x)/(length(x)+1))
  # Fit gpd to exceedances and correct their transformation
  fit = gpd.fit(x, threshold=u, npy=251, show=FALSE)
  z = mapply(
    function(a,b) fst_correct(a,b,fit$rate,fit$mle[1],fit$mle[2],u), x, xe
  )
  z
}
```

```{r}
source("../R_Scripts/EL_pickands_function.R")
t = seq(0,1,0.01)
```

```{r}
s1 = 1
s2 = 4

bll = apply(ll[,c(s1,s2)], 2, function(x) gpd_rescale(x,2))
A = A_tild(t,bll,0.95,F)

par(pty="s")
plot(
  t, A_tild(t,bll,0.85,F), xlim=c(0,1), ylim=c(0.5,1), type='l', lwd=2, col=c1,
  main=paste(sectors[s1],sectors[s2],sep="/"), ylab="A(t)"
)
lines(
  t, A_tild(t,bll,0.90,F), xlim=c(0,1), ylim=c(0.5,1), type='l', lwd=2, col=c2,
  main=paste(sectors[s1],sectors[s2],sep="/")
)
lines(
  t, A, xlim=c(0,1), ylim=c(0.5,1), type='l', lwd=2, col=c3,
  main=paste(sectors[s1],sectors[s2],sep="/")
)
lines(t,rep(1,length(t)), lty=3)
lines(t,sapply(t, function(x) max(x,1-x)), lty=3)
lines(rep(0.5,length(t[t>=0.5])), t[t>=0.5], lty=2)
points(t[which.min(A)],min(A), pch=16)
```

```{r}
s1 = 3
s2 = 12

bll = apply(ll[,c(s1,s2)], 2, function(x) gpd_rescale(x,2))
A = A_tild(t,bll,0.95,F)

par(pty="s")
plot(
  t, A_tild(t,bll,0.85,F), xlim=c(0,1), ylim=c(0.5,1), type='l', lwd=2, col=c1,
  main=paste(sectors[s1],sectors[s2],sep="/"), ylab="A(t)"
)
lines(
  t, A_tild(t,bll,0.90,F), xlim=c(0,1), ylim=c(0.5,1), type='l', lwd=2, col=c2,
  main=paste(sectors[s1],sectors[s2],sep="/")
)
lines(
  t, A, xlim=c(0,1), ylim=c(0.5,1), type='l', lwd=2, col=c3,
  main=paste(sectors[s1],sectors[s2],sep="/")
)
lines(t,rep(1,length(t)), lty=3)
lines(t,sapply(t, function(x) max(x,1-x)), lty=3)
lines(rep(0.5,length(t[t>=0.5])), t[t>=0.5], lty=2)
points(t[which.min(A)],min(A), pch=16)
```




